---
title: "covid 19"
author: "Carolina Alameda"
date: "Fall 2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

````{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```

This first plot is counting the amount of people that have caught COVID and have died or are still alive. It is divided by sex, female and male. In the second plot we read that about 60% of the females are alive and about 40% have died. You can also tell with the chart that males are dying more frequently than females almost twice as much.




  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

In this table, we can read the results of females and males out of 100 that have died or are still alive. In our previous chart we estimated that males have a higher percentage of death than females. According to this chart it is almost double when comparing female and male death. Males have a 3% while females have a 1.4%.

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The chi sq test provides the exact numbers. This test shows that 1831 were females and 1411 males were alive. Along with showing that 25 females and 44 males died.

The fisher test gives us the odds of dying for males which is 44/1411 and the odds of females dying which is 25/1831. The odds ratio is 44/1411 divided by 25/1831 is 2.3. The odds ration shows that males are 2.3x more likely to die than females.

## Conclusions

To conclude, the data shows males being more likely to die than females from covid. However the data was taken at the beginning of covid so not much information was avaliable. 



